---
title: "Practice 2.2"
author: "Euchie Jn Pierre"
date: "2022-10-02"
output: html_document
---

### This is homework is to practice using the dplyr and tidyr packages to manipulate and modify the dataset
```{r,  eval=T, include=FALSE}
#loading libraries ----
library(dplyr)
library(tidyverse)
library(knitr)
```





## **Importing the dataset using URL for reproducibility**
```{r,  eval=T}
# Preparing data----
rairuoho <- read.delim ('https://www.dipintothereef.com/uploads/3/7/3/5/37359245/rairuoho.txt')
head(rairuoho)
```





## **Manipulating data**
### Changing a name in the dataset
```{r,  eval=T, echo=TRUE}
# replacing 'nutrient' with 'enriched' within treatment
rairuoho_1 <- rairuoho %>% mutate(treatment = replace(treatment, treatment == 'nutrient', 'enriched'))
head(rairuoho_1)
```

### Reformatting a variable
```{r,  eval=T, echo=TRUE}
# reformatting 'day' as a single variable containing 6 levels and placing values into new 'length' variable.
rairuoho_1<-rairuoho_1%>%pivot_longer(day3:day8,names_to = "day", values_to = "length")
head(rairuoho_1)
```

### Uniting two variables
```{r,  eval=T,echo=TRUE}
# combining two columns (spatial1 and spatial2) into one column
rairuoho_2 <- rairuoho_1 %>%  unite("spatial coordinates", spatial1:spatial2, sep = "_")
head(rairuoho_2)
```

### Removing variables in the dataset
```{r,  eval=T, echo=TRUE}
# removing variables 'row' and 'column'
rairuoho_3 <- dplyr::select(rairuoho_2, -row, -column)
head(rairuoho_3)
```





## **The combination of the above code using the 'PIPE operator' for conciseness**
```{r,  eval=T, echo=TRUE}
# Combination of above code using the PIPE operator----
rairuoho_3_1 <- rairuoho %>% mutate(treatment = replace(treatment, treatment == 'nutrient', 'enriched'))%>%pivot_longer(day3:day8,names_to = "day", values_to = "length")%>%  unite("spatial coordinates", spatial1:spatial2, sep = "_")%>% dplyr::select(-row, -column)
head(rairuoho_3_1)
```

